---
title: "STAT 301 Group Project"
author: "Feifei Li"
date: "2026-04-09"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load library
```{r}
library(tidyverse)   
library(dplyr)       
library(ggplot2)     
library(readxl)
library(car)
```
## View the raw dataset
Raw CSV → Fix duplicates → Clean names → Handle NAs → 
Remove duplicates → Convert types → Feature engineering → 
Validation → Clean dataset 

```{r}
# import dataset
df <- read.csv("original_supply_chain.csv", stringsAsFactors = FALSE)
head(df)
```

# HANDLE Ambiguous COLUMN NAMES

```{r}
#  current column names
colnames(df)
# Problem: "Lead time" and "Lead times" are confusing, after reading the notebook in Kaggle,we need to make sure the lead_time are clearly stated 
```
# Rename the ambiguous columns

```{r}
# Rename ambiguous columns
df <- df %>%
  rename(
    Supplier_Lead_Time= `Lead.times`,           # first Lead time column
    Order_Lead_Time = `Lead.time`,           # the second lead time column
    manufacturing_lead_time = `Manufacturing.lead.time`,
    Total_Costs = `Costs`   # The last column
  )

# Verify fix
colnames(df)

```
# CHECK FOR MISSING VALUES
```{r}
# Summary of missing values by column
df %>%
  summarise_all(~ sum(is.na(.))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "na_count") %>%
  filter(na_count > 0)
# Check for missing values (handle if any)
df <- df %>%
  drop_na()  # Simple approach - remove rows with NAs

```
# Check for duplicates
```{r}
# Check duplicate rows based on all columns
duplicate_rows <- df %>% duplicated()
sum(duplicate_rows)

# View duplicate rows if any
df[duplicate_rows, ]

```
#

```{r}
# Convert categorical columns to factors
df <- df %>%
  mutate(
    `Product.type` = as.factor(`Product.type`),
    `Customer.demographics` = as.factor(`Customer.demographics`),
    `Shipping.carriers` = as.factor(`Shipping.carriers`),
    `Supplier.name` = as.factor(`Supplier.name`),
    `Location` = as.factor(`Location`),
    `Inspection.results` = as.factor(`Inspection.results`),
    `Transportation.modes` = as.factor(`Transportation.modes`),
    `Routes` = as.factor(`Routes`)
  )
```

## VERIFY NUMERICAL VARIABLE
```{r}
# Verify numeric columns
numeric_cols <- df %>% select(where(is.numeric)) %>% colnames()
print(paste("Numeric columns:", paste(numeric_cols, collapse = ", ")))
```
## OUTLIER DETECTION & HANDLING
```{r}
# Function to detect outliers using IQR method
detect_outliers <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(x < lower_bound | x > upper_bound)
}

# Apply to numeric columns and count outliers
df %>%
  select(where(is.numeric)) %>%
  summarise_all(~ sum(detect_outliers(.))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "outlier_count") %>%
  filter(outlier_count > 0)
# Visualize outliers
df %>%
  select(Price, Number.of.products.sold, Revenue.generated,Stock.levels,Defect.rates ) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(y = value)) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Outlier Detection for Key Variables")

```
## Complete ANOVA-Ready Dataset and make a new dataset named Cleaned_Supply_Chain_Data

```{r}
 # Create cleaned dataset as a new object
Cleaned_Supply_Chain_Data <- df
head(Cleaned_Supply_Chain_Data)
```
# Write the new dataset
```{r}
# Check dimensions
cat("Original dataset rows:", nrow(df), "\n")
cat("Cleaned dataset rows:", nrow(Cleaned_Supply_Chain_Data), "\n")
cat("Cleaned dataset columns:", ncol(Cleaned_Supply_Chain_Data), "\n")

# Save the cleaned dataset as a CSV file
write.csv(Cleaned_Supply_Chain_Data, "Cleaned_Supply_Chain_Data.csv", row.names = FALSE)

# Also save as RDS to preserve data types (factors, etc.)
saveRDS(Cleaned_Supply_Chain_Data, "Cleaned_Supply_Chain_Data.rds")

# Optional: Save in your working directory with a message
cat("✅ Cleaned dataset saved as 'Cleaned_Supply_Chain_Data.csv'\n")
cat("📁 Location:", getwd(), "\n")
```
# VERFIY
```{r}
Cleaned_df <- read.csv("Cleaned_Supply_Chain_Data.csv", stringsAsFactors = FALSE)
head(Cleaned_df)
cat("\n✅ Cleaned dataset successfully created and saved!\n")
```
#

```{r}
# Summary of ALL columns (numeric + categorical)
summary(Cleaned_Supply_Chain_Data)
```
# Summary on Separate summaries for better organization
```{r}
# Numeric columns summary
cat("=== NUMERIC COLUMNS SUMMARY ===\n")
numeric_cols <- Cleaned_Supply_Chain_Data %>% select(where(is.numeric))
summary(numeric_cols)

# Categorical columns summary
cat("\n=== CATEGORICAL COLUMNS SUMMARY ===\n")
categorical_cols <- Cleaned_Supply_Chain_Data %>% select(where(is.factor))
summary(categorical_cols)
```
## ANOVA TESTS
# Q1 :Research Question: Does the Product Type significantly affect the Manufacturing Costs?
```{r}
# ============================================
# QUESTION 1: One-Way ANOVA 
# Does Product Type significantly affect Manufacturing Costs?
# ============================================
 
# H_0:μ_=μ_green=μ_blue
# H_a:at least two means are different

model1 <- aov(Manufacturing.costs ~ Product.type, data = Cleaned_df)
summary(model1)
#use TukeyHSD(model) in R to do the multiple comparison
# determine which means are different.
TukeyHSD(model1)

```
# Graph
```{r}
# ============================================
# QUESTION 1: One-Way ANOVA Visualization 
# Does Product Type significantly affect Manufacturing Costs?
# ============================================
# ensure product.type is a factor
Cleaned_df$Product.type <- as.factor(Cleaned_df$Product.type)
Cleaned_df%>%
  group_by(Product.type)%>%
  summarize(AveManufacturingCosts = mean(Manufacturing.costs),
            seManufacturingCosts=sd(Manufacturing.costs)/sqrt(length(Manufacturing.costs)),
            tstar = qt(1-0.05/2, length(Manufacturing.costs)-1)
            )%>%
  ggplot(aes(x= Product.type,y=AveManufacturingCosts))+
  geom_point()+geom_errorbar(aes(ymin = AveManufacturingCosts-seManufacturingCosts*tstar, 
                                 ymax = AveManufacturingCosts +seManufacturingCosts*tstar))+
  labs(title = "Mean Manufacturing Costs by Product Type",
       x = "Product Type",
       y = "Mean Manufacturing Costs")

```
# Updated plot

# Q2 Research Question: How do Shipping Carrier and Location interact to affect Shipping Costs?

```{r}
# ============================================
# QUESTION 2: Two-Way ANOVA 
# How do Shipping Carrier and Location interact to affect Shipping Costs?
# ============================================

# The datasets are not balaced, so we need to use 
library(car)
model <-lm(Shipping.costs~ Shipping.carriers+ Location+ Shipping.carriers * Location, data= Cleaned_df)

Anova(model, type = "II")

```
## Interaction Graph
```{r}
# ============================================
# QUESTION 2: Two-Way ANOVA  VISUALIZATION FOR UNBALANCED DESIGN
# How do Shipping Carrier and Location interact to affect Shipping Costs?
# ============================================
# 
# Use $ to specify the dataframe
interaction.plot(Cleaned_df$Shipping.carriers, 
                 Cleaned_df$Location, 
                 Cleaned_df$Shipping.costs, 
                 fun = mean,
                 type = "b",
                 col = c("red", "blue", "green", "purple", "orange"),  # Different color for each location
                 lwd = 2,
                 xlab = "Shipping Carrier",
                 ylab = "Mean Shipping Cost ($)",
                 trace.label = "Location",
                 main = "Interaction Plot: Shipping Carrier × Location")

# ============================================
# Estimated Marginal Means (Adjusted for Imbalance)
# ============================================

# Run this to get the cell counts table
table(Cleaned_df$Shipping.carriers, Cleaned_df$Location)
# Method 2: With row and column totals (matches your table format)
# Create contingency table
cell_counts <- table(Cleaned_df$Shipping.carriers, Cleaned_df$Location)

# Add row totals
row_totals <- cbind(cell_counts, Total = rowSums(cell_counts))

# Add column totals
final_table <- rbind(row_totals, Total = colSums(row_totals))

# Print the table
print(final_table)
```
The final Q1&Q4 as final questions in our final 

```{r}
# ============================================
# QUESTION 1: One-Way ANOVA
# Does Product Type significantly affect Manufacturing Costs?
# Module 3 — Four Steps of ANOVA
# ============================================

# Step 1: State hypotheses (in report)
# H0: mu_cosmetics = mu_haircare = mu_skincare
# Ha: at least two means are different

# Step 2: Choose one-way ANOVA
# One categorical IV (Product.type, 3 levels), one numeric DV (Manufacturing.costs)
# ANOVA is chosen over multiple t-tests because running 3 pairwise t-tests
# at alpha=0.05 inflates Type I error — overall confidence drops from 95% to ~85.7%
# ANOVA performs one single test using F = MSB / MSW

# Ensure Product.type is a factor before modeling
Cleaned_df$Product.type <- as.factor(Cleaned_df$Product.type)

# Fit one-way ANOVA model
model1 <- aov(Manufacturing.costs ~ Product.type, data = Cleaned_df)

# Step 3: Get test statistic and p-value
summary(model1)

# Step 4: Decision
# p = 0.692 > 0.05 → fail to reject H0
# NOTE: TukeyHSD is NOT run because the omnibus F-test was not significant
# Per Module 3: post-hoc is only needed when ANOVA shows significant differences

# ============================================
# Q1 VISUALIZATION — Bar chart with 95% CI error bars
# Professor instruction: bar chart for means, NOT boxplot (boxplot = medians)
# ============================================

# Compute group means, standard errors, and 95% CI bounds per group
q1_summary <- Cleaned_df %>%
  group_by(Product.type) %>%
  summarize(
    n                     = n(),
    AveManufacturingCosts = mean(Manufacturing.costs),
    seManufacturingCosts  = sd(Manufacturing.costs) / sqrt(n()),
    tstar                 = qt(1 - 0.05 / 2, n() - 1),
    CI_low                = AveManufacturingCosts - seManufacturingCosts * tstar,
    CI_high               = AveManufacturingCosts + seManufacturingCosts * tstar
  )

# geom_col() draws bar from 0 up to the mean — shows mean magnitude visually
# geom_errorbar() adds 95% CI whiskers on top of each bar
# First geom_text() shows the exact mean dollar value above each error bar
# Second geom_text() shows sample size inside each bar (unbalanced design)
ggplot(q1_summary, aes(x = Product.type,
                        y = AveManufacturingCosts,
                        fill = Product.type)) +
  geom_col(width = 0.5, color = "black", alpha = 0.8) +
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    width = 0.15,
    linewidth = 0.8
  ) +
  geom_text(
    aes(y = CI_high,
        label = paste0("Mean=$", round(AveManufacturingCosts, 2))),
    vjust = -0.6,
    size = 3.5,
    fontface = "bold"
  ) +
  geom_text(
    aes(y = 3, label = paste0("n = ", n)),
    vjust = 0,
    size = 3.2,
    color = "white",
    fontface = "italic"
  ) +
  scale_fill_manual(values = c("cosmetics" = "#4472C4",
                                "haircare"  = "#ED7D31",
                                "skincare"  = "#70AD47")) +
  coord_cartesian(ylim = c(0, 70)) +
  labs(
    title    = "Figure 1: Mean Manufacturing Costs by Product Type",
    subtitle = "Error bars represent 95% confidence intervals",
    x        = "Product Type",
    y        = "Mean Manufacturing Costs ($)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```


