---
title: "STAT 301 Final Project"
author: "Kathryn Stoven"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(readxl)
library(tidyverse)
library(car)
library(emmeans)
```
## Read Dataset

```{r}
supply <- Cleaned_Supply_Chain_Data
names(supply)
```


# Basic data cleaning
```{r}
supply <- supply %>%
  mutate(
    Supplier.name = as.factor(Supplier.name),
    Transportation.modes = as.factor(Transportation.modes),
    Product.type = as.factor(Product.type)
  )

# Optional: verify structure
str(supply)

```

# QUESTION 3: ONE-WAY ANOVA
Research Question:
Does mean supplier lead time differ by supplier?
DV = Supplier_Lead_Time (numeric)
IV = Supplier.name (categorical)

```{r}
# -----------------------------
# 3A. Keep only needed variables
# -----------------------------
q3_data <- supply %>%
  select(Supplier.name, Supplier_Lead_Time) %>%
  drop_na()

# View sample sizes by supplier
table(q3_data$Supplier.name)

# -----------------------------
# 3B. Descriptive statistics table
# -----------------------------
q3_summary <- q3_data %>%
  group_by(Supplier.name) %>%
  summarise(
    n = n(),
    mean_lead_time = mean(Supplier_Lead_Time),
    sd_lead_time = sd(Supplier_Lead_Time)
  )

q3_summary

# -----------------------------
# 3C. Boxplot
# -----------------------------
ggplot(q3_data, aes(x = Supplier.name, y = Supplier_Lead_Time)) +
  geom_boxplot() +
  labs(
    title = "Supplier Lead Time by Supplier",
    x = "Supplier",
    y = "Supplier Lead Time"
  ) +
  theme_minimal()

# -----------------------------
# 3D. One-way ANOVA
# -----------------------------
q3_model <- aov(Supplier_Lead_Time ~ Supplier.name, data = q3_data)
summary(q3_model)

# -----------------------------
# 3E. Assumption checks
# -----------------------------

# Normality of residuals
shapiro.test(residuals(q3_model))

# Equal variance
leveneTest(Supplier_Lead_Time ~ Supplier.name, data = q3_data)

# Diagnostic plots
par(mfrow = c(2, 2))
plot(q3_model)
par(mfrow = c(1, 1))

# -----------------------------
# 3F. Post hoc test if ANOVA is significant
# -----------------------------
TukeyHSD(q3_model)

# -----------------------------
# 3G. Optional means plot with CI
# -----------------------------
q3_plot_data <- q3_data %>%
  group_by(Supplier.name) %>%
  summarise(
    mean = mean(Supplier_Lead_Time),
    sd = sd(Supplier_Lead_Time),
    n = n(),
    se = sd / sqrt(n),
    tstar = qt(0.975, df = n - 1),
    lower = mean - tstar * se,
    upper = mean + tstar * se
  )

ggplot(q3_plot_data, aes(x = Supplier.name, y = mean)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  labs(
    title = "Mean Supplier Lead Time by Supplier (95% CI)",
    x = "Supplier",
    y = "Mean Supplier Lead Time"
  ) +
  theme_minimal()
```

# QUESTION 4: TWO-WAY ANOVA WITH INTERACTION
Research Question:
Do transportation mode and product type interact to affect defect rates?
DV = Defect.rates (numeric)
IV1 = Transportation.modes (categorical)
IV2 = Product.type (categorical)
```{r}
# -----------------------------
# 4A. Keep only needed variables
# -----------------------------
q4_data <- supply %>%
  select(Transportation.modes, Product.type, Defect.rates) %>%
  drop_na()

# Check counts per cell (important for balanced vs unbalanced design)
table(q4_data$Transportation.modes, q4_data$Product.type)

# -----------------------------
# 4B. Descriptive statistics table
# -----------------------------
q4_summary <- q4_data %>%
  group_by(Transportation.modes, Product.type) %>%
  summarise(
    n = n(),
    mean_defect = mean(Defect.rates),
    sd_defect = sd(Defect.rates),
    .groups = "drop"
  )

q4_summary

# -----------------------------
# 4C. Interaction plot
# -----------------------------
interaction.plot(
  x.factor = q4_data$Transportation.modes,
  trace.factor = q4_data$Product.type,
  response = q4_data$Defect.rates,
  fun = mean,
  ylab = "Mean Defect Rate",
  xlab = "Transportation Mode",
  trace.label = "Product Type",
  main = "Interaction Plot: Transportation Mode and Product Type"
)

# -----------------------------
# 4D. ggplot interaction figure
# -----------------------------
ggplot(q4_summary,
       aes(x = Transportation.modes,
           y = mean_defect,
           group = Product.type,
           linetype = Product.type,
           shape = Product.type)) +
  geom_point(size = 3) +
  geom_line() +
  labs(
    title = "Mean Defect Rate by Transportation Mode and Product Type",
    x = "Transportation Mode",
    y = "Mean Defect Rate"
  ) +
  theme_minimal()

# -----------------------------
# 4E. Two-way ANOVA with interaction
# Recommended because design is unbalanced:
# use lm() + car::Anova(type = "II")
# -----------------------------
q4_model <- lm(Defect.rates ~ Transportation.modes * Product.type, data = q4_data)
Anova(q4_model, type = "II")

# -----------------------------
# 4F. Assumption checks
# -----------------------------
# Normality of residuals
shapiro.test(residuals(q4_model))

# Equal variance across groups
leveneTest(Defect.rates ~ Transportation.modes * Product.type, data = q4_data)

# Diagnostic plots
par(mfrow = c(2, 2))
plot(q4_model)
par(mfrow = c(1, 1))

# -----------------------------
# 4G. Post hoc comparisons
# Useful for main effects or simple effects
# -----------------------------
emmeans(q4_model, pairwise ~ Transportation.modes)
emmeans(q4_model, pairwise ~ Product.type)

# If interaction is significant, compare transportation modes within each product type
emmeans(q4_model, pairwise ~ Transportation.modes | Product.type)

# Or compare product types within each transportation mode
emmeans(q4_model, pairwise ~ Product.type | Transportation.modes)
```

